On modeling for Kerr black holes: Basis learning, QNM frequencies, and spherical-spheroidal mixing coefficients

### 2018-08-23 docs/mvfits_paper/mvf.tex
misc updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  %
  Models of black hole properties are an important component to the ongoing direct detection of gravitational waves from black hole binaries.
  %
  In particular, model based detection, and subsequent parameter estimation are directly enabled by the modeling of information related to perturbed Kerr black holes.
  %
  Here, we summarize recent methods and models for the analytically understood gravitational wave spectra (quasi-normal mode frequencies), and harmonic structure of Kerr black holes (mixing coefficients between spherical and spheroidal harmonics).
  %
  Towards the construction of these models, two algorithms are presented.
  %
  A greedy-multivariate-polynomial (GMVP) regression method and greedy-multivariate-rational (GMVR) regression method are presented for the automated modeling of polynomial and rational functions.
  %
  GMVR is a quasi-linear numerical method for interpolating rational functions. It therefore represents a solution to the problem of Runge's phenomenon.
	%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

They are the eigen-solutions of the source free linearized \ee{} (i.e. \tk{'s} equations) for a perturbed \bh{} with final mass, $\mf$, and dimensionless final spin, $\jf$.

@@

\par In the first and second lines of \eqn{hrd}, we relate the observable \gw{} polarizations, $h_+$ and $h_\times$, to the analytically understood morphology of the time domain waveform.

@@

Here, the labels $\ell$ and $m$ are eigenvalues of \tk{'s} angular equations, where units of $M=1$ (e.g. the initial mass of the \bbh{} system), and $c=1$.

@@

The consequences of this choice are discussed in reference \check{[X]}.

@@

For the so-called \textit{Phenom} models, the frequency domain multipoles, $\tilde{h}\LM(f)$, are constructed such that their high frequency behavior is consistent with \eqn{hrd} in the time domain.

@@

\par Both Phenom and \eob{} approaches directly use phenomenological models (i.e. fits) for the \qnm{} frequencies, as these fits are more computationally efficient than the underlying analytic calculations, which involve the solving of continued fraction equations.

@@

In the case of \texttt{PhenomHM} and derivative models, fits for the \qnm{} frequencies are used in the process of mapping $\tilde{h}_{22}(f)$ into other $\tilde{h}\LM(f)$ \check{[X]}.

@@

In that setting, it is demonstrated that \qnm{} frequencies correlate with the morphology of each $h\LM$ in not only \rd{}, but also merger and late inspiral, as is implied by the causal connectedness of pre and post merger.

@@

\par For models that assist tests of the \nht{}, and thereby only include precise \rd{s}, the perspective of \eqn{hrd}'s second and third lines are used to write each spherical harmonic multipole moment as

@@

\par In practice, \eqn{hlm} is computationally efficient:

@@

The calculation of each ${_{-2}S}\lmn$ involves a series solution which slowly converges for $\jf$ near unity. Therefore, it is more effective to have accurate models for $\sigma\LMlmn$, which can then be used directly to calculate $h\LM$ via \eqn{hlm}, and thereby the \gw{} polarizations via \eqn{hrd}.

@@

\par In this combined context, it is clear that the modeling of \qnm{} frequencies, $\cw\lmn$, and spherical-spheroidal mixing coefficients, $\sigma\LMlmn$, underpin a wide range of \gw{} signal models.

@@

While models for $\cw\lmn$ and $\sigma\LMlmn$ are present in many publications, there exist minor shortcomings which we wish to address here.

@@

\par For the \qnm{} frequencies, it is well known that for nearly extremal \bh{s} (i.e. $\jf \rightarrow 1$) some of the frequencies have zero-damping (i.e. $\tau\lmn \rightarrow \infty$).

@@

In the context of \aligo{} \da{}, where source parameters are estimated using routines which sample over the space of possible \bh{} masses and spins, it is useful to have accurate physical behavior in the extremal limit.

@@

\par For the modeling of $\sigma\LMlmn$, we note that the results presented in \red{[X]} are limited to cases where the azimuthal indices, $\ell$ and $\bar{\ell}$, are less than or equal to 3.

@@

As the most advanced signal models include at least $\ell$ or $\bar{\ell}$ of 3, there is use in extending prior results.

@@

In particular, it is well known that $\sigma_{43330}$ can have a significant impact on the $(4,3)$ spherical multipole.

@@

The description of \gmvp{} given here is complementary to similar algorithms used to model \qnm{} excitation amplitudes, $A\lmn$, as present in reference \red{[cite]}.

@@

As we will discuss, the \gmvr{} algorithm is an iterative approach to the (pseudo) non-linear modeling of multivariate rational functions.

@@

Both \gmvp{} and \gmvr{} are intended for use with low noise data (e.g. the results of analytic calculations), and each employs a reverse (or negative) greedy algorithm to counter over modeling.

@@

In \sec{discuss}, we review the performance of \gmvp{} and \gmvr{}, and we discuss potential applications for these methods.

@@

Within the topic of regression, linear regression has particular advantages:

@@

its matrix based formulation can be computationally efficient, and it does not require initial guesses for model parameters.

@@

Perhaps most intriguingly, the formal series expansions of smooth functions support linear and rational models (e.g. Pade approximants) having application to many datasets.

@@

Here we will develop algorithms for the linear modeling of scalar functions (real or complex valued) of many variables.

@@

With $\phi_{k}(\vec{x}_j)$ assumed to be known, the linear representation (\ceqn{lin1}) lastly defined by the set of $\mu_k$.

@@

In particular, defining $\hat{U} = \{\phi_{kj}\}$, and $\vec{f} = \{f_j\}$, then $\vec{f} = \hat{U} \; \vec{\mu}$.

@@

Equations (\ref{lin1}), and related discussion through \eqn{pinv1} illustrate the most rudimentary solution to the linear modeling problem.

@@

Second, we will build upon the \gmvp{} approach by considering models of rational functions: polynomials divided by polynomials.

@@

To consider these two approaches in a largely automated way (i.e. where the set of possible basis functions is known, but the select basis functions ultimately used are \textit{learned}), we will make use of the \textit{greedy} algorithm approach.

@@

\par A well known method for finding an approximate solution to this problem is the so-called ``greedy'' algorithm.
The premise is this:

@@

The process begins by finding the single symbol (basis vector) that yields the most accurate model (in the sense of minimizing least-squares error).

@@

In each following iteration, an all remaining symbols are added to the model one at a time, and the single additional symbol which yields the greatest increase in model accuracy is kept for the following iteration.

@@

@@

This latter point is key to the perspective of \gmvp{}: given training data thought to be related to a smooth multivariate function, it may be well approximated by a truncated series expansion in an appropriate coordinate basis.

@@

In this setting, the uncertainty of which and how many basis terms to include makes this a problem ripe for the application of \eqn{eq:linsol1} and Alg. (\ref{alg:pgreedy}) .

@@

\par In the current setting, the basis symbols required by Alg. (\ref{alg:pgreedy}) are the multinomial terms in \eqn{eq:mvt}.

@@

To counter this, we may incrementally increase, or \textit{temper}, the maximum allowed multinomial order.

@@

Recall that a term such as $x_0x_1x_3^2$ has an effective degree of 4.

@@

The combination of degree tempering with the greedy approach results in the \gmvp{} algorithm as presented in Alg. (\ref{alg:gmvp}).

@@

\par Despite the apparent universality of of \eqn{eq:mvt}, there are many cases where $K$ must be orders of magnitude greater than 1 in order for $\vec{f}$ to be accurately represented by a multinomial (historically, this is equivalent to ``Runge's phenomenon'').

@@

Of the simplest examples are the rational functions of the form

@@

where $\muf$ is the additive mean of $f(\vec{x})$, and $\sif$ is the standard deviation of $f(\vec{x})$ from its mean, and $\phi_k$ are the multinomials basis functions considered in the previous section.

@@

then the algebraic manipulation of \eqn{eq:rat1} results in

@@

At this stage, we will also explicitly consider the $j^\mathrm{th}$ samples of the domain, and so refer to $\vec{x}$ as $\vec{x}_j$.

@@

\par The key to robustly solving for $\vec{\alpha}$ lies in the iterative refinement.

@@

To this end, if we define $g^{(0)}=g$ (i.e. the training data) then, with $\hat{P} = \hat{P}( \vec{g}^{\,(n)} )$, \eqn{eq:pinv2} generalizes to

@@

  g^{(n+1)} =  \frac{ \sum_{r=0}^{R} a^{(n+1)}_r \, \phi_{r}(\vec{x}) }{ 1\;-\;  \sum_{v=1}^{V} b^{(n+1)}_v \, \phi_{v}(\vec{x}) } \; ,

@@

which is then fed back into \eqn{eq:pinv3} for further refinement.

@@

In principle, the existence of $a_r$ and $b_v$ makes the problem more complicated, as one could imagine optimizing over each symbol space independently.

@@

However, rather than independent greedy optimizations for the numerator and denominator bases symbols, \eqn{eq:rat3} suggests that the appropriate labeling of symbols (e.g. ``numerator'' or ``denominator'') may yield an effective flattening of the supposed 2D basis selection problem.

@@

Put another way, rather than two simultaneous greedy optimizations over $R$ and $V-1$ symbols (with $(V-1)R$ iterations), a single greedy process over $V+R$ symbols is performed.

@@

\par The combination of \eqns{eq:pinv3} and \eqn{eq:rat4}, along with \pgreedy{} and degree tempering, results in the \gmvr{} algorithm as presented in Alg. (\ref{alg:gmvr}).

@@

\par While not treated here, it is noted that Ref. \cite{London:2018gaq} utilizes \gmvp{} with a parameter dimension of 4.

@@

\par In the following section we apply both \gmvp{} and \gmvr{} to the modeling of noiseless (up to roundoff error) data: the \qnm{} frequencies of Kerr \bh{s}, and the inner-products of the spherical and spheroidal harmonics.
In each case, the underlying 1D parameter in each case will be the dimensionless \bh{} spin, $\jf{}$.

@@

Here, our goal is to very briefly overview the functionality of the \gmvr{} algorithm as implemented in Ref. \red{[positive]}.

@@

Similarly, we note that \gmvr{} as implemented in Ref. \red{[positive]} involves a reverse (or negative) greedy step to counter over-modeling in cases where the aforementioned $tol$ is too low.

@@

\par Let's start by applying \gmvr{} to a fiducial scalar function of the form

@@

  % f(x_0,x_1) = 50 + \frac{1.1 +0.2 x_0+0.5 x_1+x_0 x_1}{ 0.9 + x_0^2 + x_1^2 }

@@

\fig{fig:gmvrtoy}'s central bottom panel displays the distribution of fractional residuals, and demonstrates their approximately flat distribution. A gaussian fit to the fractional residuals is displayed only for comparison. In particular, a biased fit will often have residuals that are approximately gaussian; we see that this is not the case here, and that the uniformly random noise distribution is approximately recovered.

@@

When considering many noise realizations, we find that sample noise and residuals have an average correlation of $99.46\%$.

@@

Here it is demonstrated that \gmvr{} in a way that is approximately exponential, owing to the underlying analytic nature of the training data.

@@

\end{table}
%
%
\par In this example case, \gmvr{} correctly recovers the functional form of the input data. But, in general, \gmvr{} and related techniques, having no knowledge of , will attempt to model minor correlations and offsets within the training data's noise.
%
However, we have demonstrated the utility of \gmvr{} in a relatively ideal usage case where the underlying function is rational, and the training data is inly weakly contaminated with noise.
%
In the following sections, we consider cases where the functional form of the sample data is not known to be explicitly rational, but the amount of noise within the training data is negligible.

@@

Namely, we will impose a zero-damping constraint: some frequencies are real as $\jf \rightarrow 1$.

@@

appropriately linearizes the sharp behavior of each frequency near $\jf=1$, and maps $-1 \le \jf \le 1$ onto $0 \le \kappa \le 1$.

@@

For the right and left panels, simple polynomial behavior of each curve is a result of the displayed linear domain in $\kappa(\jf,\ell,m)$.

@@

In the left panel, we have scaled $1/\tau\lmn$ by factors for $m/(n+1/2)$ to place the \qnm{s} with $n=0$ and $n=1$ at approximately the same scale.

@@

\par \fig{fig:qnm}'s top left and right panels demonstrate the effect of mapping $\jf$ onto $\kappa$.

@@

In particular, it is shown that the two solution branches ($\jf>0$ and $\jf<0$) naturally form a single family of solutions when accounting for the sign of the \bh{}'s oriented spin.

@@

Concurrently, the use of $\kappa$ as a domain variable has the desirable effect of making each $\cw\lmn$ and $\tau\lmn$ approximately parabolic.

@@

We note that, in the asymptotic vicinity of $\jf=1$, the \qnm{} frequencies and decay times are known to have solutions that are asymptoticly degenerate.

@@

In allowing \eqns{eq:cw_fit_1}{eq:cw_fit_9} to extrapolate to $\jf \sim 1$, we do not explicitly account for this additional effect.

@@

\par Lastly, \fig{fig:qnm}'s bottom two panels show residual errors.

@@

\par While it is possible to enforce extremal Kerr and Schwarzschild limiting conditions for $\sigma_{\LMlmn}$, we find it effective to first use \gmvr{} to determine a functional form that works for all $\sigma_{\LMlmn}$, and then apply this form to the modeling of all $\sigma_{\LMlmn}$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  %
  Models of black hole properties play an important role in the ongoing direct detection of gravitational waves from black hole binaries.
  %
  In particular, model based gravitational wave detection, and subsequent estimation of source parameters are supported by the low level modeling of information related to perturbed Kerr black holes.
  %
  Here, we summarize recent methods and models for the analytically understood gravitational wave spectra (quasi-normal mode frequencies), and harmonic structure of Kerr black holes (mixing coefficients between spherical and spheroidal harmonics).
  %
  Towards the construction of these models, two algorithms are presented.
  %
  A greedy-multivariate-polynomial (GMVP) regression method and greedy-multivariate-rational (GMVR) regression method are presented for the automated modeling of polynomial and rational functions respectively.
  %
  GMVR is a quasi-linear numerical method for interpolating rational functions. It therefore represents a solution to the problem of Runge's phenomenon.
	%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

They are the eigen-solutions of the source free linearized \ee{} (i.e. \tk{'s} equations \cite{PhysRevLett.29.1114}) for a perturbed \bh{} with final mass, $\mf$, and dimensionless final spin, $\jf$.

@@

\par In the first and second lines of \eqn{hrd}, we relate the observable \gw{} polarizations, $h_+$ and $h_\times$, to the analytically understood morphology of the time domain \rd{} waveform.

@@

Here, the labels $\ell$ and $m$ are polar and azimuthal eigenvalues of \tk{'s} angular equations, where units of $M_i=1$ (e.g. the initial mass of the \bbh{} system), and $c=1$.

@@

The consequences of this choice, in particular mixing between spherical and spheroidal harmonics, are discussed in reference \cite{London:2014cma,London:2018gaq}.

@@

For the so-called \textit{Phenom} models, the frequency domain multipoles, $\tilde{h}\LM(f)$, are constructed such that their high frequency behavior is consistent with \eqn{hrd} in the time domain \cite{Hannam:2013oca,London:2017bcn,Khan:2015jqa,Schmidt:2014iyl}.

@@

\par Both Phenom and \eob{} approaches directly use phenomenological models (i.e. fits) for the \qnm{} frequencies, as these fits are more computationally efficient than the underlying analytic calculations, which involve the solving of continued fraction equations \cite{Leaver85}.

@@

In the case of \texttt{PhenomHM} and derivative models, fits for the \qnm{} frequencies are used in the process of mapping $\tilde{h}_{22}(f)$ into other $\tilde{h}\LM(f)$ \cite{London:2017bcn}.

@@

In that setting, it is demonstrated that \qnm{} frequencies correlate with the morphology of each $h\LM$ in not only \rd{}, but also merger and late inspiral, as is implied by the source's causal connectedness pre and post merger.

@@

\par For models that assist tests of the \nht{} (e.g. \cite{Berti:2005ys,London:2018gaq,Carullo:2018sfu}), and thereby only include precise \rd{s}, the perspective of \eqn{hrd}'s second and third lines are used to write each spherical harmonic multipole moment as

@@

\par In practice, using \eqn{hlm} is computationally efficient:

@@

Whereas the calculation of each ${_{-2}S}\lmn$ involves a series solution which slowly converges for $\jf$ near unity, the calculation of each ${_{-2}Y}\LM$ is achieved using closed form expressions.
It is therefore efficient to avoid convergence issues by using accurate models for $\sigma\LMlmn$, which can then be used directly to calculate $h\LM$ via \eqn{hlm}, and thereby the \gw{} polarizations via \eqn{hrd}.

@@

\par In this combined context, it is clear that the modeling of \qnm{} frequencies, $\cw\lmn$, and spherical-spheroidal mixing coefficients, $\sigma\LMlmn$, are relevant for a range of \gw{} signal models.

@@

While models for $\cw\lmn$ and $\sigma\LMlmn$ are present in the literature, there exist minor shortcomings which we wish to address here \cite{Berti:2005ys,Berti:2014fga}.

@@

\par For the \qnm{} frequencies, it is well known that for nearly extremal \bh{s} (i.e. $\jf \rightarrow 1$) some of the frequencies have zero-damping (i.e. $\tau\lmn \rightarrow \infty$) \cite{Zimmerman:2015trm}.

@@

In the context of \aligo{} \da{}, where source parameters are estimated using routines which sample over the space of possible \bh{} masses and spins, it is useful to have accurate physical behavior in the extremal limit \cite{TheLIGOScientific:2016wfe}.

@@

\par For the modeling of $\sigma\LMlmn$, we note that the results presented in \cite{Berti:2014fga} are limited to cases where the azimuthal indices, $\ell$ and $\bar{\ell}$, are less than or equal to 3.

@@

As the most advanced signal models include at least $\ell$ or $\bar{\ell}$ of 3, there is use in extending current results.

@@

In particular, it is well known that $\sigma_{43330}$ can have a significant impact on the $(4,3)$ spherical multipole (See \ceqn{hlm}).

@@

The description of \gmvp{} given here is complementary to similar algorithms used to model \qnm{} excitation amplitudes, $A\lmn$, as present in reference \cite{London:2018gaq,London:2014cma}.

@@

As we will discuss, the \gmvr{} algorithm is an iterative approach to the (pseudo) linear modeling of multivariate rational functions, wherein iterations of linear inversions are used to refine the ultimately non-linear model.

@@

\par In the rudimentary form presented here, both \gmvp{} and \gmvr{} are intended for use with low noise data (e.g. the results of analytic calculations), and each employs a reverse (or negative) greedy algorithm to counter over modeling \cite{Field:2011mf,Caudill:2011kv}.

@@

In \sec{discuss}, we review the performance of \gmvp{} and \gmvr{}, and we discuss current and potential applications for these methods.

@@

Within the topic of regression, linear regression has particular advantages.

@@

Its matrix based formulation can be computationally efficient, and it does not require initial guesses for model parameters.

@@

Perhaps most intriguingly, the formal series expansions of smooth functions support linear and rational models (e.g. Pad\'e approximants) having application to many datasets.

@@

Here we will develop algorithms for the linear (polynomial and rational) modeling of scalar functions (real or complex valued) of many variables.

@@

With $\phi_{k}(\vec{x}_j)$ assumed to be known, the linear representation (namely \ceqn{lin1}) is lastly defined the set of $\mu_k$.

@@

In particular, defining $\hat{U} = \{\phi_{kj}\}$, and $\vec{f} = \{f_j\}$, then $\vec{f} = \hat{U} \; \vec{\mu}$, implies that

@@

\par Equations (\ref{lin1}), and related discussion through \eqn{pinv1} illustrate the most rudimentary solution to the linear modeling problem.

@@

Second, we will build upon the \gmvp{} approach by considering models of rational functions (polynomials divided by polynomials).

@@

To consider these two approaches in a largely automated way (i.e. where the set of possible basis functions is known, but the select basis functions ultimately used are \textit{learned}), we will make use of the \textit{greedy} algorithm approach \cite{Field:2011mf}.

@@

\par A well known method for finding an approximate solution to this problem is the so-called ``greedy'' algorithm:

@@

The process begins by finding the single symbol (basis vector) that yields the most accurate model (e.g. in the sense of minimizing the least-squares error).

@@

In each new iteration, an all remaining symbols are added to the model one at a time, resulting in many trial models.
When considering all trial models, the single models symbol which yields the greatest increase in model accuracy is kept for the subsequent greedy iteration.

@@

\par The very similar ``negative'' greedy algorithm removes model symbols until representation error increases beyond a specified threshold.

@@

\par This latter point is key to the perspective of \gmvp{}:
given training data thought to be related to a smooth multivariate function, it may in many cases be well approximated by a truncated series expansion in an appropriate coordinate basis.

@@

In this setting, the uncertainty of which and how many basis terms to include makes this a problem ripe for the application of linear modeling driven by a greedy process, namely \eqn{eq:linsol1} and Alg. (\ref{alg:pgreedy}) .

@@

\par In this setting, the basis symbols required by Alg. (\ref{alg:pgreedy}) are the multinomial terms in \eqn{eq:mvt}.

@@

To counter this, we may incrementally increase, or \textit{temper}, the maximum allowed multinomial degree.

@@

For example, when iterating through allowed degrees, if the current maximum degree is 3, then degree 4 terms, such as $x_0x_1x_3^2$, will not be considered within the space of model symbols.

@@

The degree tempering process halts when increasing the maximum allowed degree has no significant effect on model representation error.
\par The combination of degree tempering with the greedy approach results in the \gmvp{} algorithm as presented in Alg. (\ref{alg:gmvp}).

@@

\par Despite the apparent universality of of \eqn{eq:mvt}, there are many cases where $K$ must be orders of magnitude greater than 1 in order for $\vec{f}$ to be accurately represented by a multinomial.

@@

Worst, in cases where the underlying dataset is best described by a rational function, no polynomial order will yield satisfactory results (i.e. the well known ``Runge's phenomenon'').
\par Of the simplest examples are rational functions of the form

@@

where $\muf$ is the additive mean of $f(\vec{x})$, and $\sif$ is the standard deviation of $f(\vec{x})$, and $\phi_k$ are the multinomials basis functions considered in the previous section.

@@

then algebraic manipulation of \eqn{eq:rat1} allows

@@

At this stage, we will also explicitly consider the $j^\mathrm{th}$ samples of the domain, and so refer to (e.g.) $\vec{x}$ as $\vec{x}_j$.

@@

\par The key to robustly solving for $\vec{\alpha}$ lies in iterative refinement \cite{Press:1992:NRC:148286}.

@@

To this end, if we define $g^{(0)}=g$ (i.e. $g$ is the training data) then, with $\hat{P} = \hat{P}( \vec{g}^{\,(n)} )$, then \eqn{eq:pinv2} generalizes to

@@

  g^{(n+1)} =  \frac{ \sum_{r=0}^{R} a^{(n+1)}_r \, \phi_{r}(\vec{x}) }{ 1\;-\;  \sum_{v=1}^{V} b^{(n+1)}_v \, \phi_{v}(\vec{x}) } \; .

@@

Subsequently, $g^{(n+1)}$ is then fed back into \eqn{eq:pinv3} for further refinement.

@@

In principle, the existence of $a_r$ and $b_v$ makes the problem more complicated, as one might imagine optimizing over each symbol space independently.

@@

However, rather than independent greedy optimizations for the numerator and denominator bases symbols, \eqn{eq:rat3} suggests that the appropriate labeling of symbols (e.g. ``numerator'' or ``denominator'') may yield an effective flattening of the supposed 2D symbol selection problem.

@@

Put another way, rather than two simultaneous greedy optimizations over $R+1$ and $V$ symbols (with $(R+1)V$ iterations), a single greedy process over $V+R+1$ symbols is performed.

@@

\par The combination of \eqn{eq:pinv3} and \eqn{eq:rat4}, along with \pgreedy{} and degree tempering, results in the \gmvr{} algorithm as presented in Alg. (\ref{alg:gmvr}).

@@

@@

Both \gmvp{} and \gmvr{} are publicly available on Github through the \texttt{positive} repository \cite{lionel_london_2018_1402516}.

@@

Here, our goal is to very briefly overview the functionality of the \gmvr{} algorithm as implemented in Ref. \cite{lionel_london_2018_1402516}.

@@

Similarly, we note that \gmvr{} as implemented in Ref. \cite{lionel_london_2018_1402516} involves a negative greedy phase to counter over-modeling in cases where the aforementioned $tol$ is too low.

@@

\par Let us now consider the application of \gmvr{} to a fiducial scalar function of the form

@@

@@

\fig{fig:gmvrtoy}'s central bottom panel displays the distribution of fractional residuals, and demonstrates their approximately flat distribution.

@@

A gaussian fit to the fractional residuals is displayed for comparison.
In particular, despite the uniform nature of the underlying noise distribution, a biased fit will often have residuals that are approximately gaussian.
We see that this is not the case here, and that the uniformly random noise distribution is approximately recovered.
Specifically, when considering many noise realizations, we find that sample noise and residuals have an average correlation of $99.46\%$.

@@

Here it is demonstrated that \gmvr{} converges in a way that is approximately exponential, owing to the underlying analytic nature of the training data.

@@

\end{table}
\par In this rudimentary example case, \gmvr{} correctly recovers the functional form of the input data. But, in general, \gmvr{} and related techniques, having no knowledge of the underlying noise distribution, will attempt to model minor correlations and offsets within the training data's noise.
However, we have demonstrated the utility of \gmvr{} in a relatively ideal usage case where the underlying function is rational, and the training data is inly weakly contaminated with noise.
\par In the following sections, we consider still ideal cases, where the functional form of the sample data is not known to be explicitly polynomial or rational, but the amount of noise within the training data is negligible.

@@

Namely, we will impose a zero-damping constraint: some frequencies are real as $\jf \rightarrow 1$ \cite{Zimmerman:2015trm}.

@@

appropriately linearizes the sharp behavior of each frequency near $\jf=1$ while also mapping $-1 \le \jf \le 1$ onto $0 \le \kappa \le 1$.

@@

For the top right and top left panels, the simple polynomial behavior of each curve is a result of the displayed linear domain in $\kappa(\jf,\ell,m)$.

@@

In the top left panel, we have scaled $1/\tau\lmn$ by factors for $m/(n+1/2)$ to place the \qnm{s} with $n=0$ and $n=1$ at approximately the same scale.

@@

\par \fig{fig:qnm}'s top left and right panels' upper axis demonstrate the effect of mapping $\jf$ onto $\kappa$.

@@

In particular, it is shown that the two solution branches ($\jf>0$ and $\jf<0$) naturally form a single family of solutions when accounting for the sign of the \bh{}'s oriented spin \cite{Husa:2007hp}.

@@

Concurrently, the use of $\kappa$ as a domain variable has the desirable effect of making each $\cw\lmn$ and $\tau\lmn$ approximately quarctic.

@@

We note that, in the asymptotic vicinity of $\jf=1$, the \qnm{} frequencies and decay times are known to have solutions that are asymptoticly degenerate \cite{Zimmerman:2015rua}.

@@

In allowing \eqns{eq:cw_fit_1}{eq:cw_fit_9} to extrapolate to $\jf=1$, we do not explicitly account for this additional effect.

@@

\par Lastly, \fig{fig:qnm}'s bottom two panels show fractional residual errors.

@@

\par While it is possible to enforce extremal Kerr and Schwarzschild limiting conditions for $\sigma_{\LMlmn}$, we find it effective to first use \gmvr{} to determine a functional form that works for \textit{all} $\sigma_{\LMlmn}$, and then apply this form to the modeling of all $\sigma_{\LMlmn}$.
